---
title: "Cheese over Space and Animals"
description: Scatterplot analysis of cheese origins over nations
author: Jazzmin Duncan
date: February 17, 2025
format: html
execute: 
  warning: false
  message: false
---

## Introduction

For this analysis, I'll be investigating the relationships between dairy (and non-dairy) products across the world. Specifically, looking at the different animals used for milk/cheese, I hope to draw conclusions about potential eating habits or traditions across the world. 

### Data

The data being used for this plot was originally collected by cheese.com, an archive and database containing an incredible amount of information about cheeses. [Used for Tidy Tuesday](https://github.com/rfordatascience/tidytuesday/tree/a95bcce0d7e09590519ba1ff98512fbf10f443af/data/2024/2024-06-04) in June of 2024 to investigate the relationships between fat content, cheese type, texture, flavor, and aroma -- this data contains 1,187 rows of different cheeses, and 19 columns specifying the details of said cheeses. For today, the relevant columns are cheese (name), milk, country, and color.

Here is a sneak peak at the actual data. 

```{r}
# Loading in data and displaying first 10 rows

cheeses <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2024/2024-06-04/cheeses.csv')
head(cheeses, 10)
```

## Analysis

```{r, echo=FALSE}
# Loading in libraries for processing string data and data verbs

library(dplyr)
library(stringr)

# Removes unneeded columns
# Removes empty values from relevant columns
# Sorts by same country and milk type
# Counts number of cheeses and makes new column
# Only includes countries with more than one cheese
cheese_important <- cheeses |> 
  select(cheese, milk, country, color) |>
  filter(!is.na(milk), !str_detect(country, ","), !str_detect(milk, ",")) |>
  group_by(country, milk) |> 
  summarize(num_cheeses = n(), .groups = "drop") |>
  mutate(num_cheese = as.integer(num_cheeses)) |>
  filter(num_cheese > 1)
```

### Plotting

```{r, fig.width=10, fig.height=8, echo=FALSE}

library(ggplot2)

ggplot(cheese_important, aes(x = reorder(country, -(num_cheeses)), y = num_cheeses, fill = milk)) +
  geom_bar(stat = "identity") +
  facet_wrap(~milk, scales = "free_y") +
  coord_flip() +
  labs(title = "Magnitude of Cheese Types (milk used) by Country",
       x = "Country",
       y = "Number of Different Cheeses of Type in Countries",
       fill = "Milk") +
  theme(
    
    # Making text readable (would be too large/overlapping without)
    text = element_text(size = 12),
    axis.text.y = element_text(size = 8),
    axis.text.x = element_text(size = 10),
    strip.text = element_text(size = 10, face = "bold"))
    

```

This is a faceted bar chart demonstrating different eating habits across nations as it pertains to milk-cheese production. It uses the name of each country as the x-axis, with the total amount of a certain type of cheese (which has been counted per country) as the y-axis. The fill color and faceting are by the specific animal the milk came from. 

### Observations and Conclusions

We can see that cows are substantially more popular in the global cheese industry, while other milks, like water buffalo, go relatively unnoticed. It would be interesting to see this data overlayed with/compared to species/location/time data to see if cheese popularity is in any way connected to seasonal migrations or changes in availability. You can also see countries in multiple different groups demonstrating both the breadth and nicheness of their cheeses, Italy being particularly noticeable in this regard -- as they take up two entire facets on their own (water buffalo and buffalo). 
