---
title: "Cheese over Space and Animals"
description: Scatterplot analysis of cheese origins over nations
author: Jazzmin Duncan
date: February 17, 2025
format: html
execute: 
  warning: false
  message: false
---

The data being used for this plot was originally collected by cheese.com, an archive and database containing an incredible amount of information about cheeses. Used for Tidy Tuesday in June of 2024 to investigate the relationships between fat content, cheese type, texture, flavor, and aroma -- this data contains 1,187 rows of different cheeses, and 19 columns specifying the details of said cheeses. For today, the relevant columns are cheese (name), milk, country, and color.


```{r}
cheeses <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2024/2024-06-04/cheeses.csv')
cheeses
```

```{r, echo=FALSE}
library(dplyr)
library(stringr)
cheeses <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2024/2024-06-04/cheeses.csv')

cheese_important <- cheeses |> 
  select(cheese, milk, country, color) |>
  filter(!is.na(milk), !str_detect(country, ","), !str_detect(milk, ","))

cheese_summary <- cheese_important |> 
  group_by(country, milk) |> 
  summarize(num_cheeses = n(), .groups = "drop") |>
  mutate(num_cheese = as.integer(num_cheeses)) |>
  filter(num_cheese > 1)
```

```{r, fig.width=10, fig.height=8, echo=FALSE}
library(ggplot2)

ggplot(cheese_summary, aes(x = reorder(country, -(num_cheeses)), y = num_cheeses, fill = milk)) +
  geom_bar(stat = "identity") +
  facet_wrap(~milk, scales = "free_y") +
  coord_flip() +
  labs(title = "Magnitude of Cheese Types (milk used) by Country",
       x = "Country",
       y = "Number of Different Cheeses of Type in Countries",
       fill = "Milk") +
  theme(
    text = element_text(size = 12),
    axis.text.y = element_text(size = 8),
    axis.text.x = element_text(size = 10),
    strip.text = element_text(size = 10, face = "bold"))

```
This is a faceted bar chart demonstrating different eating habits across nations as it pertains to cheese. We can see that cows are substantially more centered in the dairy industry, while other milks, like water buffalo, go relatively unnoticed. You can also see countries in multiple different groups demonstrating both the breadth and nicheness of their cheeses, Italy being particularly noticible in this regard. 

TidyTusedy dataset: https://github.com/rfordatascience/tidytuesday/tree/a95bcce0d7e09590519ba1ff98512fbf10f443af/data/2024/2024-06-04

