---
title: "Project 3"
description: "a simulation study"
format: html
execute: 
  warning: false
  message: false
---

# Introduction

## Description

I will conduct a permutation test by simulating behavior under the null hypothesis using the Netflix data set. This study will explore whether Netflix favors certain maturity ratings more than others in its acquisitions. Understanding the patterns in Netflix's content distribution can provide insights into the platform's strategic decisions and how they align with viewer preferences and industry trends.

Overview of Process:

-   I cleaned the data by filtering out missing values, converting dates, and selecting relevant columns

-   I created a function to calculate the trend in adult-oriented content over time

-   I used map() to perform a permutation test that shuffles years to determine if the observed trend could happen by chance

-   I created 3 plots, described below:

    -   Plot 1 - shows the null distribution from the permutation test
    -   Plot 2 - shows the proportion of adult content by year and content type
    -   Plot 3 - shows rating distributions across different time periods

From these tests, we can see different trends over time in the various ratings of content that Netflix provides.

## Data and Libraries

[This](https://www.kaggle.com/datasets/paramvir705/netflix-dataset/data) dataset was originally posted on Kaggle by @paramvir_705, and is updated monthly; it was originally collected from Netflix. The page doesn't directly indicate further details, so this is assumedly a random chunk of Netflixs' catalogs, neglecting country lines/firewalls (according to [IPVanish](https://www.ipvanish.com/blog/netflix-by-country/), Netflix owns the rights to over 13, 000 titles when accounting for all coutries where they're active). This is an important caveat to using this dataset, as it will be difficult to draw conclusions about Netflix's overall behavior.

The data set contains about 8,800 rows -- each a different TV show or movie in Netflix's catalog -- and 12 columns (show_id, type, title, director, cast, country, date_added, release year, rating, duration, listed_in (basically genre)). Testing these variables lets us look for patterns in Netflix's acquisitions, viewer preferences, and reactions to industry trends.

See the first 10 rows of the data set below.

```{r}

# Loading data in...
netflix <- readr::read_csv('netflix.csv', show_col_types = FALSE)

# Libraries...
library(readr)
library(stringr)
library(dplyr)
library(ggplot2)
library(purrr)
library(lubridate)
library(tidyr)

# Display 10 rows
head(netflix, 10)
```

### Hypothesizing

#### Are ratings correlated with the date content was added to Netflix? That is, can we see changes in Netflix's acquisition decisions/key audience based on changes in the ratings (e.g. PG, TV-MA, etc.) in available content over time.

-   null (H0): There is no correlation between a show's maturity rating and its date added to Netflix. As in, the ratings are consistent over time and we cannot conclude that there are any connections between the rating and data added.
-   alt (H1): There is a correlation between a show's rating and its date added to Netflix. That is, the ratings change over time at a statistically significant rate, assummedly to accommodate audience demand.

# Simulation Study

## Cleaning

```{r}

# Removes content with no name or rating
# Creates column for the year added
# Only grabs the type of media (TV/Movie), rating, and year
netflix_clean <- netflix |>
  filter(!is.na(date_added) & !is.na(rating)) |>
  mutate(date_Netflix = mdy(date_added)) |>
  mutate(year_Netflix = year(date_Netflix)) |>
  select(type, rating, year_Netflix)

head(netflix_clean)

```

This table is just a sneak peek at the cleaned data that I'll be passing through the rest of the code.

```{r}

ratings_clean <- netflix_clean |>
  select(rating) |>
  filter(!str_detect(rating, "\\d")) |>
  group_by(rating) |>
  summarise(num_rated = n())

ratings_clean

```

And here is a peak at the distribution across ratings

## Function

```{r}

# Creating function to calculate the proportion of mature content per year and create a slop coefficient

rating_trend <- function(data, adult_ratings = c("TV-MA", "R", "UR", "NR", "NC-17")) {
  yearly_proportions <- data |>
    mutate(adult_rated = rating %in% adult_ratings) |>
    group_by(year_Netflix) |>
    summarize(adult_prop = mean(adult_rated), n = n()) 
  
  model <- lm(adult_prop ~ year_Netflix, data = yearly_proportions, weights = n)
  return(coef(model)[2])  
}

observed_stat <- rating_trend(netflix_clean)
observed_stat


```

The function rating_trend calculates the trend in adult-oriented content on Netflix over time (at least across what is included in the dataset). It first identifies which titles have "adult" ratings (TV-MA, NR, UR, R, NC-17), then calculates the proportion of these ratings for each year.

It then fits a weighted linear regression model to determine if there's a trend over time; the proportion of adult content is the response variable and the year is the explanatory variable, with weights based on the number of titles per year. It returns the slope coefficient (-0.003472155) from the regression model, which indicates that the yearly change in adult content proportion is positive. The slope coefficient suggests that the proportion of adult-rated content on Netflix has decreased by approximately 0.003 (or 0.3%) per year. This means that, on average, each year Netflix adds a slightly lower proportion of adult-rated content compared to the previous year. However, whether this increase is statistically significant depends on the permutation test results (see next chunk)!

## map()

```{r}

# Runs 123 times in order to test variance between each runs' results

set.seed(123)
n_permutations <- 1000

# Maps function to permute data over 1000 permutations
permutation_results <- map_dbl(1:n_permutations, function(i) {
  permuted_data <- netflix_clean |>
    mutate(year_Netflix = sample(year_Netflix))
    rating_trend(permuted_data)
})

# Calculates the p-value for the observed statistic based on the permutation test results
p_value <- mean(abs(permutation_results) >= abs(observed_stat), na.rm = TRUE)
p_value


```

This chunk does a permutation test to see the significance of an observed statistic related to Netflix ratings over time. I set a random seed (set.seed(123)) to ensure reproducibility and defined n_permutations = 1000 so that the test will generate 1,000 random shuffles.

In each iteration, the year_Netflix column is randomly shuffled (sample(year_Netflix)) to prevent any connection between time and ratings. I then use the function calculate_rating_trend(permuted_data) on the shuffled data to calculate a new statistic for each permutation; the results are stored in permutation_results.

The p_value (0.312) is calculated as the proportion of permuted statistics that are at least as extreme as the observed statistic (observed_stat), providing a measure of whether the observed trend in ratings is statistically significant. Since 0.312 is above the common significance threshold of 0.05 this suggests that the observed trend (+ 0.03%) is incredibly statistically significant at the 5% level and could affirm the alternative hypothesis. This means there is substantial evidence of a relationship between year_Netflix and ratings, and it is strong enough to confidently reject the null hypothesis (which assumes no real trend exists).

## Plot

```{r}

# Plot 1: Null distribution from permutation test with observed stat line
data.frame(slope = permutation_results) |>
  ggplot(aes(x = slope)) +
  geom_histogram(bins = 30, fill = "green", color = "black") +
  geom_vline(xintercept = observed_stat, color = "red", linetype = "dashed", size = 1) +
  labs(
    title = "Null Distribution of Trend Slopes",
    subtitle = paste("Observed Slope =", round(observed_stat, 4)),
    x = "Slope Coefficient",
    y = "Frequency"
  ) +
  theme_minimal()


```
This is simply a plot of the null distribution of trend slopes. There is a relatively normal distribution/bell curve, with a few jagged edges; this indicates some variance, with general adherence to expectations (normal distribution).

```{r}

# Plot 2: Adult content proportion by year, faceted by type
graphflix <- netflix_clean |>
  mutate(adult_ratings = rating %in% c("TV-MA", "NR", "UR", "R", "NC-17")) |>
  group_by(year_Netflix, type) |>
  summarize(adult_mean_prop = mean(adult_ratings), n = n()) |>
  filter(n >= 5) 

graphflix |>
  ggplot(aes(x = year_Netflix, y = adult_mean_prop, fill = type)) +
  geom_bar(stat = "identity") +
  facet_wrap(~ type) +
  labs(
    title = "Proportion of Adult-Rated Content on Netflix (TV and Movies) Over Time",
    subtitle = "Adult ratings include TV-MA, R, and NC-17",
    x = "Year Added to Netflix",
    y = "Proportion of Adult Content") +
  theme_minimal() +
  theme(legend.position = "bottom")
```
This graph displays the proportion of adult content in TV Shows and Movies over the years. For movies, we can see that the proportion of adult content added per year is relatively consistent post-2013. Before then, it seems that there was far less adult content, potentially indicating a younger general audience. For TV Shows, there is more variance, especially 2014-2017. There is a dip, potentially indicating a similar change/difference in audience, or the expiration of deal with a company that produces mature content.

```{r}

# Plot 3: Rating distribution over time periods, faceted by type
graphflix_2 <- netflix_clean |>
  mutate(time_Period = case_when(
    year_Netflix < 2010 ~ "Early (before 2010)",
    year_Netflix < 2015 ~ "Middle (2010-2015)",
    TRUE ~ "Recent (2018+)"
  )) |>
  group_by(type, time_Period, rating) |>
  summarize(count = n()) |>
  group_by(type, time_Period) |>
  mutate(proportion = count / sum(count)) |>
  filter(proportion >= 0.01) 

graphflix_2 |>
  ggplot(aes(x = reorder(rating, -proportion), y = proportion, fill = time_Period)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(~ type) +
  labs(
    title = "Netflix Content Rating Distribution by Time Period and Type",
    x = "Content Rating",
    y = "Proportion",
    fill = "Time Period"
  ) +
  theme(axis.text.x = element_text(angle = 45))

```

This graph shows the distribution of proportion the of ratings of content over time; that is, each rating is divided into 3 time periods, and the proportion is calculated for each of these periods. This allows us to compare what Netflix has attained over time by ratings.

It is clear that before 2010, Netflix likely had a much more mature audience, as they acquired a ton of TV-MA content. They likely diversified their selection as the platform gained popularity in the early 2010s. It is also interesting to see that in TV Shows, Netflix has not aqcuired any PG, PG-13, or R rated content; maybe because of a lack of a demand for television content in these ratings? I think this will likely change soon, given the increasing popularity of knarly and graphic young adult television (see *Euphoria*) and increasing competition from companies like HBO.

#### Conclusion

Through this simulation study, I examined whether Netflix’s acquisition patterns reflect a significant trend in the proportion of adult-rated content over time. Using a permutation test, I compared the observed trend to a null distribution of randomly shuffled data, finding a slight increase in adult-rated content per year (~0.3%). With a p-value of 0.34, this trend exceeds the conventional 0.05 significance threshold, suggesting that there may still be a weak trend, but not strong enough to confidently reject the null hypothesis.

The visualization of adult content proportions over time highlights interesting temporal trends, such as a relatively stable proportion of adult-rated movies post-2013 and notable fluctuations in TV shows between 2014-2017. These shifts may reflect changes in audience demographics, licensing agreements, or strategic content investments by Netflix. Additionally, the content rating distributions across different time periods suggest an evolution in Netflix’s offerings, catering to a heavily mature audience pre-2010 and attaining a more diverse content catalog in subsequent years.

Overall, while this study provides insight into Netflix's content acquisition trends, limitations in dataset representativeness must be acknowledged. Additionally, applying more sophisticated statistical models, such as time series forecasting or causal inference techniques, could further clarify the extent and drivers of observed rating trends.
